\textbf{The urgency of the discussed issue.} To implement the government program <<Socio-economic development of the Arctic zone of the Russian Federation for the period until 2020>> approved by government decision \cite{PostRF} need to implement the strategy \cite{Strategi}. According to which the integration of the Arctic zone with the main regions of Russia is envisaged through: the exploitation of hydrocarbons, non-ferrous and precious metals; the formation of modern transport-logistics nodes and the core network of roads; development, reconstruction and modernization of the airport network.

This will erequire the maintenance, newly built and reconstructed roads and airfields during winter. The duration of which in some areas is more than 140 days a year. The most complex and important arrangements for the maintenance of road surfaces, for various purposes, aimed at the destruction of and removal of snow-ice formations (SIF). There are several ways of dealing with the SIF: chemical-mechanical; frictional; heat; mechanical. The latter method allows you to destroy and remove SIF from pavements without causing harm to the environment and to save on chemical reagents, fuel; to maintain the integrity of the roadway. It establishes mechanically the leader in the development and design of new workers on road cars.

However, there is a niche in which this method is not very effective, namely the removal of persistent snow-ice formations (PSIF). This is due to their physico-mechanical properties: density $\rho = 0,6 \div 0,9 \slantfrac{\text{g}}{\text{sm}^3}$.; the limit of compressive strength $\sigma = 2,5 \div 2,8$ MPa.; the thickness of the layer $h \le 100$ mm.; the temperature of the test medium $-2{}^\circ C\div-10{}^\circ C$. Existing working bodies or are not adapted to their destruction, or make it not effective. To improve performance and reduce energy consumption when you remove DSIF proposed the use of disk cutting tool \cite{GanjaDRI, GanjaPSLO, WorkOrgan}. However, with the use of disk cutting tool, the problem of creation of highly effective working bodies, at the design stage need to know about force paremeters, the value of which depends on many factors. Such as: cutting speed; the geometrical parameters of the instrument; ambient temperature and destructible material; value of the radius of curvature of the working edge. 

\textbf{The main aim of the study:} preparation of the instrumental base for determining the dependence of the force parameters, namely force of resistance to cutting persistent snow-ice formations on factors such as the radius of curvature of the working edge of the disk cutting tool and the step of cutting. The work is a continuation of a series of experimental researches of processes of interaction of disc cutting tools with PSIF, for a number of years, held in Siberian Federal University.

\textbf{The methods used in the study:} When solving the problem we used a comprehensive approach including: scientific analysis and generalization of the experience of earlier studies; calibration of the measuring transducer on a special test bench; mathematical and statistical processing of the results of the calibration.

\textbf{The results:} The calibration coefficients of the voltage conversion from the measuring transducer to the value of the force for each component were obtained. Conclusions are made about the correctness of the selection of the material of the tezometric element and the correctness of the paste on of the tezoresistors. The hypothesis on exclusion of mutual influence of measured components against each other is confirmed.